---
title: "Geography 176A"
author: "[Shannon Sweeney]https://github.com/Shannonfsweeney"
subtitle: 'Lab 03: COVID-19 Pandemic'
output:
  html_document: html_document
  theme: journal
---
install.packages("tidyverse")
library("tidyverse")
install.packages("knitr")
install.packages("readxl")
install.packages("zoo")
install.packages("dplyr")
library(zoo)
install.packages("magrittr")
```{r load-packages, include=FALSE}
library(dplyr)
library(magrittr)
library(knitr)
```
### Question 1 
```{r, message=FALSE, warning=FALSE}
library(tidyverse)

url = 'https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv'
raw_covid_data = read_csv(url)
```
```{r}
question1 = (raw_covid_data %>% 
  select(date, fips ,county, state, cases) %>% 
  filter(state == 'California') %>% 
  group_by(state) %>% 
  mutate(daily_cases = cases - lag(cases, default = 0)) %>% 
  ungroup())
```
### Cumulative California Cases
```{r}
cum_cases = (question1 %>% 
  select(date,fips, county, cases) %>% 
  filter(date == max(date)) %>%
  arrange(-(cases)) %>% 
  head(5))
```



### California Counties table
```{r}
knitr::kable(cum_cases, caption = "California Counties with the most COVID-19 cases" )
```



### code for second half of question  1
```{r}
pt2question1 = (raw_covid_data %>% 
  select(date, state, county, fips, cases) %>% 
  filter(state == 'California') %>%
  filter(date >= ('2020-08-02')) %>% 
  group_by(county)) %>% 
  mutate(total_cases=max(cases)-min(cases)) %>%
  arrange(-(total_cases)) %>% 
  select(date, county, total_cases, fips) %>% 
   filter(date == max(date)) %>% 
  head(5)
```
```{r}
pt2question1 %>% 
 filter(date == max(date)) %>% 
  head(5)
```
```{r}
knitr::kable(pt2question1, caption = "California Counties with the most COVID-19 cases within the  past 14 days" )
```
```{r}
library(readr)
landdata_states <- read_csv("../data/landdata-states.csv")
```
```{r}
library(readxl)
PopulationEstimates <- read_excel("../data/PopulationEstimates.xls", 
    skip = 2)
```
```{r}
p1 = PopulationEstimates %>%
  select(fips = "FIPStxt",state = "State", pop2019 = "POP_ESTIMATE_2019") %>%
  filter(state == "CA") 
```

```{r}
popcases = cum_cases %>% full_join(p1) 
```
```{r}
percapita_cumcases = (popcases %>% 
  head(5) %>% 
  mutate(per_capita_case_ratio = 100000* (cases/pop2019)))
```
###final table for question one 
```{r}
knitr::kable(percapita_cumcases, caption = " California Counties with the most cumulative cases of Covid, including cases per 100,000 people")
```
### All of California Couties with the most COVID-19 cases are in Southern California.
```{r}
table2data= (pt2question1 %>% full_join(p1))
```
```{r}
fixedtable2data = table2data %>% 
  mutate(per_capita_case_ratio = 100000* (total_cases/pop2019)) %>% 
  head(5)
```
```{r}
knitr::kable(fixedtable2data, caption = "COVID 19 cases per 100,000 people in the California Counties with the most cases in the last 14 days")
```
### All five of the Counties with the most COVID-19 cases in the past 2 weeks exceed the watchlist limit, 25 cases per 100,000 within the last 14 days.

### Question 2
### California
```{r}
California = raw_covid_data %>%
  filter(state == 'California') %>%
  group_by(date) %>%
  summarise("State_Cases" = sum(cases)) %>% 
  mutate(daily_cases = State_Cases - lag(State_Cases))
```
```{r}
California %>% 
 ggplot(aes(x = date)) +
  geom_line(aes(y = daily_cases), col = "coral1", size = 1) 
```
###New York
```{r}
newYork = raw_covid_data %>%
  filter(state == 'New York') %>%
  group_by(date) %>%
  summarise("State_Cases" = sum(cases)) %>% 
  mutate(daily_cases = State_Cases - lag(State_Cases)) %>% 
  select(date, daily_cases, State_Cases)
```
### I was having trouble getting the roll mean function to work.
```{r}
newYork %>% 
 ggplot(aes(x = date)) +
  geom_line(aes(y = daily_cases), col = "coral1", size = 1) 
```

### Louisiana
```{r}
Louisiana = raw_covid_data %>%
  filter(state == 'Louisiana') %>%
  group_by(date) %>%
  summarise("State_Cases" = sum(cases)) %>% 
  mutate(daily_cases = State_Cases - lag(State_Cases))
```
```{r}
Louisiana %>% 
 ggplot(aes(x = date)) +
  geom_line(aes(y = daily_cases), col = "coral1", size = 1) 
```
### Florida
```{r}
Florida = raw_covid_data %>%
  filter(state == 'Florida') %>%
  group_by(date) %>%
  summarise("State_Cases" = sum(cases)) %>% 
  mutate(daily_cases = State_Cases - lag(State_Cases))
```
```{r}
Florida %>% 
 ggplot(aes(x = date)) +
  geom_line(aes(y = daily_cases), col = "coral1", size = 1) 
```

